### 2009-02-19 doc/storage_system/arc-storage-documentation.tex
adding the skeleton of the User's manual renaming "arc1-storage" to "arc-storage"

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC1 storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

%%%%---- The title ----
{\centering \textsc{\Large Documentation of the ARC1 storage system}\Large \par}
\vspace*{0.5cm}

@@

The ARC1 storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.  The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.). For each storage node one of the services of the ARC1 storage system is needed to manage it and to integrate it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC1 storage system. The main services of the storage system are the following (see Figure~\ref{fig:services}):
\begin{itemize}
    \item the \textbf{A-Hash} service, which is a currently centralized, later distributed database;
    \item the \textbf{Librarian} service, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
    \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
    \item the \textbf{Shepherd} service, which provides a simple interface for storing files on storage nodes.
\end{itemize}

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc1-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC1 storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
\end{figure}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-namespace.pdf}}}
\caption{\label{fig:namespace}Example of the hierarchy of the global namespace} }
\end{figure}

@@

The current version of the ARC1 storage system has no implemented solution for security issues. Designing and implementing a powerful, flexible, fine-grained solution for authentication and authorization is very important. Currently we have plans how to do that, and the design of the storage services makes it easy to integrate the chosen solution into the system.

@@

\section{Downloading a file} % (fold)
\label{sec:downloading_a_file}
\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-downloading.pdf}}}
\caption{\label{fig:downloading}Downloading a file} }
\end{figure}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-uploading.pdf}}}
\caption{\label{fig:uploading}Uploading a file} }
\end{figure}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc1-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
\begin{enumerate}
    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
    \item The Bartender asks the Librarian to traverse the LN and return the list of parent collections of this file.
    \item The Librarian returns the data.
    \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection. If the file has more parent collections (hard links) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
    \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
\end{enumerate}

@@

The Bartender provides an easy to use interface of the ARC1 storage system to the users. You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}. The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}. A \emph{stat} call gives all the information about a file or collection, and you can move (or hardlink) collections and files within the namespace with \emph{move}. You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.

@@

There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC1 storage namespace into the local filesystem namespace, and use it with the commands of the operating system.

@@

To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC1 storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages. It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC1 storage (Section~\ref{sub:data_model}). The files here would be referenced by a path which is local in the namespace of the third-party storage. The interface could be something like this:
\begin{description}
    \item[get(getRequest)] returns list of (\emph{requestID}, \emph{getResponseData})

>>>>

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Zsombor Nagy},
  pdftitle = {Documentation of the ARC storage system},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

%%%%---- The title ----
{\centering \textsc{\Large Documentation of the ARC storage system}\Large \par}
\vspace*{0.5cm}

@@

The ARC storage system is a distributed system for storing replicated \emph{file}s on several file storage nodes and manage them in a global namespace.  The files can be grouped into \emph{collection}s (a concept very similar to directories in the common file systems), and a collection can contain sub-collections and sub-sub-collections in any depth. There is a dedicated \emph{root collection} to gather all collections to the global namespace. This hierarchy of collections and files can be referenced using \emph{Logical Name}s (\emph{LN}s). The users can use this global namespace  as they were using a local filesystem. Files can be transfered by multiple transfer protocols, and the client side tools hide this from the user. The replicas of the files are stored on different storage nodes. A storage node here is a network-accessible computer having storage space to share, and one or more data transfer service running (e.g.~HTTP(S), FTP(S), GridFTP, ByteIO\footnote{OGSA ByteIO Working Group (BYTEIO-WG), \url{https://forge.gridforum.org/projects/byteio-wg/}}, etc.). For each storage node one of the services of the ARC storage system is needed to manage it and to integrate it into the system. And there are plans to provide a way to integrate third-party storage solutions into the namespace of the ARC storage system. The main services of the storage system are the following (see Figure~\ref{fig:services}):
\begin{itemize}
    \item the \textbf{A-Hash} service, which is a currently centralized, later distributed database;
    \item the \textbf{Librarian} service, which stores the metadata and hierarchy of collections and files, the location of replicas, and health data of the Shepherd services, using the A-Hash as database;
    \item the \textbf{Bartender} service, which provides a high-level interface for the users and for other services;
    \item the \textbf{Shepherd} service, which provides a simple interface for storing files on storage nodes.
\end{itemize}

\begin{figure}[ht]
\centering{{\scalebox{0.9}{\includegraphics{arc-storage-services.pdf}}}
\caption{\label{fig:services}The components of the ARC storage: the \textbf{A-H}ash service, the \textbf{L}ibrarian service, the \textbf{B}artender service and the \textbf{S}hepherd service.}}
\end{figure}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc-storage-namespace.pdf}}}
\caption{\label{fig:namespace}Example of the hierarchy of the global namespace} }
\end{figure}

@@

The current version of the ARC storage system has no implemented solution for security issues. Designing and implementing a powerful, flexible, fine-grained solution for authentication and authorization is very important. Currently we have plans how to do that, and the design of the storage services makes it easy to integrate the chosen solution into the system.

@@

\section{Downloading a file} % (fold)
\label{sec:downloading_a_file}
\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc-storage-downloading.pdf}}}
\caption{\label{fig:downloading}Downloading a file} }
\end{figure}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc-storage-uploading.pdf}}}
\caption{\label{fig:uploading}Uploading a file} }
\end{figure}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.7}{\includegraphics{arc-storage-removing.pdf}}}
\caption{\label{fig:removing}Removing a file} }
\end{figure}
\begin{enumerate}
    \item If we want to remove a file, we should connect to a Bartender with the LN of the file we want to remove.
    \item The Bartender asks the Librarian to traverse the LN and return the list of parent collections of this file.
    \item The Librarian returns the data.
    \item If the file has only parent, it asks the Librarian to remove the file entry itself, and the entry from the parent collection. If the file has more parent collections (hard links) the Bartender only removes the entry from the parent collection and removes the parent from the list of parent  collections of this file.
    \item Next time the Shepherd does its periodic check, it asks the Librarian about each of its stored replicas, and finds out that one of them no longer exists, so it removes the replica from the storage node.
\end{enumerate}

@@

The Bartender provides an easy to use interface of the ARC storage system to the users. You can put, get and delete files using their logical names (\emph{LN}s) with the \emph{putFile}, \emph{getFile} and \emph{delFile} methods, create, remove and list collections with \emph{makeCollection}, \emph{unmakeCollection} and \emph{list}. The metadata of a file or collection (e.g.~whether the collection is closed, number of needed replicas, access policies) can be changed with \emph{modify}. A \emph{stat} call gives all the information about a file or collection, and you can move (or hardlink) collections and files within the namespace with \emph{move}. You can upload an entirely new replica to a file (e.g.~if the file lost all its replicas, or when a Shepherd service offers its replica for replications) with \emph{addReplica}.

@@

There are plans to create more sophisticated CLI and GUI tools and to create a FUSE\footnote{Filesystem in Userspace, \url{http://fuse.sourceforge.net/}} module and Windows Shell Extensions\footnote{\url{http://msdn.microsoft.com/en-us/magazine/cc188741.aspx}} to be able to mount the ARC storage namespace into the local filesystem namespace, and use it with the commands of the operating system.

@@

To integrate existing files on a third-party storage system to our namespace thus make them accessible through the interface of the ARC storage, we need a service which provides a common interface to the Bartender services, and hides the details of accessing the different third-party storages. It should translate the method calls, the \emph{get}s, \emph{put}s and \emph{remove}s and the \emph{ACL}\footnote{Access Control Lists, policies} modifications, and try to create transform metadata according to the data model of the ARC storage (Section~\ref{sub:data_model}). The files here would be referenced by a path which is local in the namespace of the third-party storage. The interface could be something like this:
\begin{description}
    \item[get(getRequest)] returns list of (\emph{requestID}, \emph{getResponseData})
